### 2022-11-02 signal_processing_elsevier/submission_ijf/cover_letter.tex
add first ijf submission

\documentclass[1p]{elsarticle}

\pagenumbering{gobble}
\makeatletter
\def\ps@pprintTitle{%
  \let\@oddhead\@empty
  \let\@evenhead\@empty
  \def\@oddfoot{\reset@font\hfil\thepage\hfil}
  \let\@evenfoot\@oddfoot
}
\makeatother

%\usepackage{lineno,hyperref}
\usepackage{hyperref}
%\usepackage[utf8]{inputenc}
%\usepackage[english]{babel}
%\usepackage{authblk}
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}
\usepackage[OT1]{fontenc}

\usepackage{amssymb,amsmath,amsthm}
%\usepackage[pdftex]{graphicx}
%\usepackage{footnote}
%\usepackage{url}
%\usepackage{times,dsfont}
%\usepackage{fancyhdr}
%\usepackage{geometry}
%\pagestyle{fancy}
%\modulolinenumbers[5]

%\journal{Journal of \LaTeX\ Templates}

\begin{document}

\begin{frontmatter}

\title{HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series}

\author{Etienne DAVID}

%% or include affiliations in footnotes:
\author{Jean BELLOT}

\author{Sylvain LE CORFF} 
\end{frontmatter}

\vspace{0.5cm}
Dear Editors,\vspace{0.5cm}

Please find attached our manuscript ”HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series”, which we
would like to submit for publication as a research article in your journal Elsevier International Journal of Forecasting.

\medskip

The fashion industry has produced many complex time series with seasonalities, trends and influences. To date, it has never really been studied by the signal processing community because of the lack of relevant data. In this work, we  aim at tackling this limitation, by translating the history of fashion trends to time series and proposing a new forecasting model able to accurately forecast them. First, by tracking thousands of fashion trends on social media with state-of-the-art computer vision approaches, we create a first fashion dataset gathering 10000 weekly fashion time series. As influence dynamics are the key of emerging trend detection, we associate with each time series an external weak signal representing behaviors of influencers. Secondly, to leverage such a complex and rich dataset, we propose a new hybrid forecasting model. Our approach combines per-time-series parametric models with seasonal components and a global recurrent neural network to include sporadic external signals. This hybrid model provides state-of-the-art results on the proposed fashion dataset, on the weekly time series of the M4 competition, and illustrates the benefit of the contribution of external weak signals.

\medskip

We would like to emphasize that contrary to some recent new forecasting studies, the fashion dataset studied in this paper will be publicly provided\footnote[1]{\url{http://files.heuritech.com/raw_files/f1_fashion_dataset.tar.xz}} upon publication of this work.

\medskip

%Firstly, based on complexe vision algorithms, we are able to track hundreds of fashion items on pictures shared on social media and draw the history of thousands of fashion trends. As the fashion sphere is highly driven by influence dynamics, we also define external signal representing behaviours of social media influencers for each trends, we call them ”weak signals”. With this paper, we provide publicly a sample of 10000 fashion time series and their linked ”weak signals”. ii) Secondly, as the previous sequences hide complexe dynamics and external signals, we introduce a new hybrid framework to accurately forecast them. This hybrid model combines per-time-series parametric models with seasonal components and a global recurrent neural network to include sporadic external signals. Our apporach provides state-of-the-art results on the proposed fashion dataset and also on the weekly time series dataset of the M4 competition.

We are looking forward to hearing from you.

\medskip

\indent Sincerely.

\medskip

Etienne David.

\end{document}