### 2023-05-16 tmlr_submission/hermes_paper_revision/answers_reviews.tex
add answers_reviews.tex in the tlmr revision dir

\documentclass[10pt]{article} % For LaTeX2e
\usepackage{tmlr}
% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{tmlr}
% To de-anonymize and remove mentions to TMLR (for example for posting to preprint servers), instead use the following:
%\usepackage[preprint]{tmlr}

% Optional math commands from https://github.com/goodfeli/dlbook_notation.
\input{math_commands.tex}

\usepackage{hyperref}
\usepackage{url}
\usepackage{amssymb,amsmath,amsthm}
\usepackage[pdftex]{graphicx}

\title{HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series}

% Authors must not appear in the submitted version. They should be hidden
% as long as the tmlr package is used without the [accepted] or [preprint] options.
% Non-anonymous submissions will be rejected without review.

\author{\name \'Etienne David \email etienne.david@heuritech.com \\
      \addr SAMOVAR, Télécom SudParis,\\
      Institut Polytechnique de Paris, 91120 Palaiseau, France
      \AND
      \name Jean Bellot \email jean.bellot@heuritech.com \\
      \addr Heuritech, \\
      6 Rue de Braque, 75003 Paris
      \AND
      \name Sylvain Le Corff \email sylvain.lecorff@gmail.com\\
      \addr LPSM, \\
      Sorbonne Université, UMR CNRS 8001, 75005, Paris
      }

% The \author macro works with any number of authors. Use \AND 
% to separate the names and addresses of multiple authors.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\def\month{01}  % Insert correct month for camera-ready version
\def\year{2023} % Insert correct year for camera-ready version
\def\openreview{\url{https://openreview.net/forum?id=XXXX}} % Insert correct link to OpenReview for camera-ready version

\begin{document}

\maketitle

We would like to thank the reviewers and the associate editor for their constructive feedback, and their appreciation of the paper. We have carefully considered all the comments listed by the reviewers and we proposed a revision of the paper accordingly. You can find below an overview of the major modifications of the manuscript followed by detailed responses to the reviewers.

\section*{Major changes}

\section*{Reviewer 1}

\begin{itemize}
	\item {\em ``The paper does not provide a clear motivation or literature review for the choice of per-time-series models.''} \medskip
	\textbf{Author response:} \\

	\item {\em ``The paper could benefit from a more detailed discussion of the limitations of the proposed approach and potential avenues for future work.''} \medskip
	\textbf{Author response:} \\

	\item {\em ``The paper could also benefit from a more thorough comparison with existing methods, including a discussion of their strengths and weaknesses in relation to the proposed approach.''} \medskip
	\textbf{Author response:} \\

	\item {\em ``The paper should provide a clear motivation or literature review for the choice of per-time-series models.''} \medskip
	\textbf{Author response:} \\

	\item {\em ``The paper should explain why exponential smoothing and TBATS are chosen as the predictors, and how they compare to other possible choices such as ARIMA or HMM.''} \medskip
	\textbf{Author response:} \\

	\item {\em ``The paper should compare with more existing methods.''} \medskip
	\textbf{Author response:} \\

	\item {\em ``The paper should explain how HERMES handles different types of time series, such as stationary, non-stationary, seasonal, or sporadic.''} \medskip
	\textbf{Author response:} \\

	\item {\em ``The paper should discuss the potential positive and negative impacts of the proposed model and dataset on the society and the environment.''} \medskip
	\textbf{Author response:} \\

	\item {\em ``The paper should address the privacy and consent issues of using social media images and data to construct the fashion dataset and the weak signals.''} \medskip
	\textbf{Author response:} \\

	\item {\em ``The paper should explain how the data is collected, processed, and anonymized, and what are the ethical and legal implications of this process.''} \medskip
	\textbf{Author response:} \\

	\item {\em ``The paper should also acknowledge the potential risks or biases of using social media data to represent fashion trends, such as cultural diversity, inclusivity, or sustainability.''} \medskip
	\textbf{Author response:} \\	
\end{itemize}

\section*{Reviewer 2}

\begin{itemize}
	\item {\em ``''} \medskip
	\textbf{Author response:} \\

	\item {\em ``''} \medskip
	\textbf{Author response:} \\

	\item {\em ``''} \medskip
	\textbf{Author response:} \\

	\item {\em ``''} \medskip
	\textbf{Author response:} \\

	\item {\em ``''} \medskip
	\textbf{Author response:} \\		
\end{itemize}

\subsection*{Reviewer 3}

\begin{itemize}
	\item {\em ``''} \medskip
	\textbf{Author response:} \\

	\item {\em ``''} \medskip
	\textbf{Author response:} \\

	\item {\em ``''} \medskip
	\textbf{Author response:} \\

	\item {\em ``''} \medskip
	\textbf{Author response:} \\

	\item {\em ``''} \medskip
	\textbf{Author response:} \\		
\end{itemize}

\end{document}