### 2023-06-05 tmlr_submission/hermes_paper_revision/answers_reviewer_1.tex
Update answers_reviewer_1.tex

\documentclass[10pt]{article} % For LaTeX2e
\usepackage{tmlr}
% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{tmlr}
% To de-anonymize and remove mentions to TMLR (for example for posting to preprint servers), instead use the following:
%\usepackage[preprint]{tmlr}

% Optional math commands from https://github.com/goodfeli/dlbook_notation.
\input{math_commands.tex}

\usepackage{hyperref}
\usepackage{url}
\usepackage{amssymb,amsmath,amsthm}
\usepackage[pdftex]{graphicx}

\title{HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series}

% Authors must not appear in the submitted version. They should be hidden
% as long as the tmlr package is used without the [accepted] or [preprint] options.
% Non-anonymous submissions will be rejected without review.

\author{\name \'Etienne David \email etienne.david@heuritech.com \\
      \addr SAMOVAR, Télécom SudParis,\\
      Institut Polytechnique de Paris, 91120 Palaiseau, France
      \AND
      \name Jean Bellot \email jean.bellot@heuritech.com \\
      \addr Heuritech, \\
      6 Rue de Braque, 75003 Paris
      \AND
      \name Sylvain Le Corff \email sylvain.lecorff@gmail.com\\
      \addr LPSM, \\
      Sorbonne Université, UMR CNRS 8001, 75005, Paris
      }

% The \author macro works with any number of authors. Use \AND 
% to separate the names and addresses of multiple authors.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\def\month{01}  % Insert correct month for camera-ready version
\def\year{2023} % Insert correct year for camera-ready version
\def\openreview{\url{https://openreview.net/forum?id=XXXX}} % Insert correct link to OpenReview for camera-ready version

\begin{document}

\maketitle

We would like to thank the reviewer for the appreciation of the paper and the constructive feedback. We have carefully considered all comments and we will propose a revision of the paper accordingly when all reviews will be published. In the meantime, you can find below detailed responses and an overview of some of the  modifications that will be added in the revised manuscript.

\section*{Detailed responses}

\subsection*{Requested changes:}

\begin{itemize}

	\item {\em ``The paper should provide a clear motivation or literature review for the choice of per-time-series models.''} \medskip

	\textbf{Author response:} We thank the referee for this remark and we acknowledge that we did not discuss enough the choice of per-time-series models in the paper. This comment has led to additional justifications and discussions, in particular we discuss the use of state space models. We highlight that the main limitation for this choice is the computational time required to train the per-time-series models. In addition, in the experiment section, a third HERMES variation is added using Thetam as the per-times-series model. Table~\ref{tab:metricresults} presents the new version of Table 2 of the paper with the new HERMES candidate that we call \textit{hermes-thetam}. Changes between the old and the new version are written in orange. We believe that the three HERMES approaches using different per-time-series models highlight the flexibility of the hybrid framework.\\

	\item {\em ``The paper should explain why exponential smoothing and TBATS are chosen as the predictors, and how they compare to other possible choices such as ARIMA or HMM.''} \medskip

	\textbf{Author response:} We thank the referee for this remark.  The main limitation is the computational time required to train these models, as the dataset contains thousands of time series. For instance, training a HMM with the EM algorithm, or training ARIMA on thousands of time series would increase significantly the computational cost. Concerning TBATS, Exponential smoothing and Thetam, correctly fitting them with existing Python packages is possible in an reasonable time, from a few minutes for exponential smoothing models to a couple of hours for TBATS. We believe that our experiment section, in particular with the additional simulations, highlights the flexibility of the hybrid approach and the numerical results motivate the use of TBATS eventhough other models could be used. To clarify this point in the paper, several justifications and comments were added in the text.\\

	\item {\em ``The paper should compare with more existing methods.''} \medskip

	\textbf{Author response:} We agree with this suggestion and this comment has led to the addition of 2 new methods as benchmarks. First, the Prophet method introduced in \citet{Taylor2017} is trained and evaluated on the fashion dataset using the Python package  \texttt{prophet}. We show that it leads to poor results in comparison to our model on the whole fashion dataset, and a comparable accuracy on the use case where only 100 time series are accessible. Secondly, the recent DeepAR method introduced in \citet{salinas2020} is added  and trained with the Python package \texttt{Gluonts} \citep{Alexandrov2020}. This method shows striking results on the fashion dataset and is a strong competitor of the proposed hybrid approach. As an illustration, Table~\ref{tab:metricresults} presents the new version of the Table 2 of the paper with the two new models: Prophet and DeepAR. Changes between the old and the new version are written in orange.\\

	\item {\em ``The paper should explain how HERMES handles different types of time series, such as stationary, non-stationary, seasonal, or sporadic.''} \medskip

	\textbf{Author response:} We thank the referee for this remark that led to a new part in the appendix where 4 sub-samples of the Fashion dataset are built and forecasted separately. The first one called \textit{disrupting} time series gathers the most difficult sequences to forecast. The second one called \textit{stable} time series defines a sample of fashion time series that are stable from a year to another. The third one called \textit{seasonal} time series gathers sequences with a huge sasonality. Finally, the last group represents the most sporadic and noisy time series of the fashion time series. We added the predictions of each model on these sub-samples and extended the analysis of the hybrid approach, the fashion dataset and the use of external signals. Table~\ref{tab:fashionsubsample} shows the results of each method on the 4 sub-samples and it will be added in appendix as well as a section describing more precisely how the 4 samples are built.\\
\end{itemize}

\begin{table}
  \caption{Results summary on the 10000ts Fashion dataset. For each metric, the average on all our time series is computed. For approaches using neural networks, 10 models are trained with different seeds. The mean and the standard deviation of the 10 results are displayed.}
  \centering
  \begin{tabular}{l||lllll|lllll}
   &&\multicolumn{3}{c}{\textbf{MASE $\downarrow$}} &&& \multicolumn{3}{c}{\textbf{ACCURACY $\uparrow$}}&\\
    &&  \textit{mean}  && \textit{std} &&&  \textit{mean}  && \textit{std}& \\
	 \hline
	 &&&&&&&&&&\\
     \textit{snaive} && 0.881 && - &&& 0.357 && - &\\
     \textcolor{orange}{\textit{thetam}} && \textcolor{orange}{0.845} && \textcolor{orange}{-} &&& \textcolor{orange}{0.463} && \textcolor{orange}{-}\\
     \textit{arima} && 0.826 && -&&& 0.464 && - & \\
     \textit{ets} && 0.807 && -&&& 0.449 && - & \\
     \textcolor{orange}{\textit{prophet}} && \textcolor{orange}{0.786} && \textcolor{orange}{-} &&& \textcolor{orange}{0.485} && \textcolor{orange}{-}\\
     \textit{stlm} && 0.770 && -&&& 0.482 && - & \\
     \textit{hermes-ets-ws} && 0.769 && 0.005 &&& 0.501 && 0.007 &\\
     \textcolor{orange}{\textit{hermes-thetam}} && \textcolor{orange}{0.764} && \textcolor{orange}{0.003} &&& \textcolor{orange}{0.497} && \textcolor{orange}{0.005}\\
     \textcolor{orange}{\textit{hermes-thetam-ws}} && \textcolor{orange}{0.760} && \textcolor{orange}{0.004} &&& \textcolor{orange}{\textbf{0.520}} && \textcolor{orange}{0.010}\\
     \textit{hermes-ets} && 0.758 && 0.001 &&& 0.490 && 0.006 &\\
     \textcolor{orange}{\textit{deepar}} && \textcolor{orange}{0.752} && \textcolor{orange}{0.018} &&& \textcolor{orange}{0.459} && \textcolor{orange}{0.015}\\
     \textit{tbats} && 0.745 && -&&& 0.453 && - & \\
     \textit{lstm-ws} && 0.728 && 0.004 &&& 0.500 && 0.008 &\\
     \textit{lstm} && 0.724 && 0.003 &&& 0.498 && 0.007 &\\
     \textit{hermes-tbats} && 0.715 && 0.002 &&& 0.488 && 0.008 &\\
     \textbf{\textit{hermes-tbats-ws}} && \textbf{0.712} && 0.004 &&& 0.510 && 0.005 &\\
  \end{tabular}
\label{tab:metricresults}
\end{table}

\begin{table}
  \caption{Results summary on 4 differents sub-sample of  Fashion time series: i) disrupting time series, ii) stable time series, iii) seasonal time series and iv) noisy time series.}\vspace{0.5cm} 
 \centering
 \resizebox{0.8\textwidth}{!}{
  \begin{tabular}{l||llll}
    \multicolumn{4}{c}{\textit{disrupting} time series}\\
    &&\multicolumn{2}{c}{\textbf{MASE}} \\
    &&  \textit{mean}  & \textit{std}  \\
	 \hline
	 &&& \\
     \textit{snaive} && 1.455 & -\\ 
	 \textit{thetam} && 1.314 & -\\ 
	 \textit{ets} && 1.27 & -\\ 
	 \textit{arima} && 1.256 & -\\ 
	 \textit{tbats} && 1.229 & -\\ 
	 \textit{hermes-thetam} && 1.209 & 0.005\\ 
	 \textit{hermes-ets} && 1.202 & 0.007\\ 
	 \textit{stlm} && 1.198 & -\\ 
	 \textit{hermes-tbats} && 1.195 & 0.01\\ 
	 \textit{prophet} && 1.193 & -\\ 
	 \textit{deepar} && 1.18 & 0.03\\ 
	 \textit{lstm} && 1.15 & 0.01\\ 
	 \textit{hermes-thetam-ws} && 1.145 & 0.019\\ 
	 \textit{hermes-ets-ws} && 1.131 & 0.024\\ 
	 \textit{hermes-tbats-ws} && 1.092 & 0.007\\ 
	 \textit{lstm-ws} && 1.086 & 0.009\\  \vspace{0.5cm}\\
  \end{tabular}\hspace{1cm}
  \begin{tabular}{l||llll}
   \multicolumn{4}{c}{\textit{stable} time series}\\
   &&\multicolumn{2}{c}{\textbf{MASE}} \\
    &&  \textit{mean}  & \textit{std}  \\
	\hline
	 &&& \\
     \textit{prophet} && 0.629 & -\\ 
	 \textit{thetam} && 0.615 & -\\ 
	 \textit{ets} && 0.611 & -\\ 
	 \textit{arima} && 0.565 & -\\ 
	 \textit{hermes-ets-ws} && 0.555 & 0.007\\ 
	 \textit{snaive} && 0.536 & -\\  
	 \textit{deepar} && 0.531 & 0.024\\ 
 	 \textit{hermes-thetam-ws} && 0.522 & 0.004\\ 
	 \textit{hermes-ets} && 0.518 & 0.002\\ 
	 \textit{stlm} && 0.513 & -\\ 
	 \textit{hermes-thetam} && 0.508 & 0.005\\ 
	 \textit{tbats} && 0.501 & -\\ 
	 \textit{lstm-ws} && 0.492 & 0.007\\ 
	 \textit{hermes-tbats-ws} && 0.477 & 0.008\\ 
	 \textit{lstm} && 0.47 & 0.004\\ 
	 \textit{hermes-tbats} && 0.451 & 0.002\\  \vspace{0.5cm}\\
  \end{tabular}
 }
 \resizebox{0.8\textwidth}{!}{ 
  \begin{tabular}{l||llll}
    \multicolumn{4}{c}{\textit{seasonal} time series}\\
    &&\multicolumn{2}{c}{\textbf{MASE}} \\
    &&  \textit{mean}  & \textit{std}  \\
	 \hline
	 &&& \\
     \textit{snaive} && 0.895 & -\\ 
	 \textit{ets} && 0.895 & -\\ 
	 \textit{prophet} && 0.851 & -\\ 
	 \textit{deepar} && 0.836 & 0.035\\ 
	 \textit{lstm-ws} && 0.829 & 0.014\\ 
	 \textit{thetam} && 0.826 & -\\ 
	 \textit{lstm} && 0.823 & 0.013\\ 
	 \textit{hermes-thetam} && 0.815 & 0.008\\ 
	 \textit{tbats} && 0.81 & -\\ 
	 \textit{hermes-ets-ws} && 0.809 & 0.01\\ 
	 \textit{hermes-thetam-ws} && 0.808 & 0.008\\ 
	 \textit{arima} && 0.805 & -\\ 
	 \textit{stlm} && 0.786 & -\\ 
	 \textit{hermes-ets} && 0.785 & 0.003\\ 
	 \textit{hermes-tbats-ws} && 0.777 & 0.012\\ 
	 \textit{hermes-tbats} && 0.772 & 0.003\\ 
  \end{tabular}\hspace{1cm}
  \begin{tabular}{l||llll}
   \multicolumn{4}{c}{\textit{noisy} time series}\\
   &&\multicolumn{2}{c}{\textbf{MASE}} \\
    &&  \textit{mean}  & \textit{std}  \\
	\hline
	 &&& \\
     \textit{snaive} && 0.842 & -\\ 
	 \textit{hermes-ets-ws} && 0.739 & 0.005\\ 
	 \textit{hermes-ets} && 0.726 & 0.002\\ 
	 \textit{ets} && 0.721 & -\\ 
	 \textit{thetam} && 0.717 & -\\ 
	 \textit{stlm} && 0.715 & -\\ 
	 \textit{prophet} && 0.698 & -\\ 
	 \textit{arima} && 0.696 & -\\ 
	 \textit{hermes-thetam-ws} && 0.672 & 0.003\\ 
	 \textit{hermes-thetam} && 0.669 & 0.001\\ 
	 \textit{deepar} && 0.661 & 0.007\\ 
	 \textit{hermes-tbats-ws} && 0.647 & 0.006\\ 
	 \textit{tbats} && 0.646 & -\\ 
	 \textit{hermes-tbats} && 0.644 & 0.003\\ 
	 \textit{lstm-ws} && 0.637 & 0.004\\ 
	 \textit{lstm} && 0.636 & 0.003\\
  \end{tabular}
 }
 \label{tab:fashionsubsample}
\end{table}

\subsection*{Broader Impact Concerns:}

\begin{itemize}
	\item {\em ``The paper should discuss the potential positive and negative impacts of the proposed model and dataset on the society and the environment.''} \medskip

	\textbf{Author response:} We agree with the suggestion of the referee and this point is now developed in the discussion section. We discuss the impact of training deep architectures and the computational cost of the proposed approach. However,we highlight that the aim of this work is to show that it is possible to better forecast and anticipate consumers behaviours in the retail and fashion industry. With a better anticipation of consumer needs, companies could better adjust their production, reduce massive wastes due to their overstocks and finally offset the environmental cost of this work.  \\

	\item {\em ``The paper should address the privacy and consent issues of using social media images and data to construct the fashion dataset and the weak signals.''} \medskip

	\textbf{Author response:} We thank the referee for this remark and several information concerning this issue were added in Section 2. The whole creation of the dataset respects the privacy and data protection regulations, including the General Data Protection Regulation. Concerning the image recovery on social media, only public accounts have been analysed and no potential private information were used, saved or revealed during the whole process. Concerning the time series, all the trends names have been anonymized and only macro information such as the geolocalisation, the type of cloth and the gender are revealed publicly. \\

	\item {\em ``The paper should explain how the data is collected, processed, and anonymized, and what are the ethical and legal implications of this process.''} \medskip

	\textbf{Author response:} We agree with the referee and  details were added  in  Section 2. \\

	\item {\em ``The paper should also acknowledge the potential risks or biases of using social media data to represent fashion trends, such as cultural diversity, inclusivity, or sustainability.''} \medskip

	\textbf{Author response:} We totally agree with the referee. First of all, an important part of this work was to design a correct normalization process to remove the most visible biases brought by social media. Then, by proposing time series splitted by geo zone, we have tried to propose an accurate representation of the diversity of behaviours that exists in each region. We acknowledge that the proposed geozones may seem large. A future work could be to built trends at country level. Finally, by proposing a global model to forecast all the fashion trends, it is possible that rare patterns may not be correctly learned by the proposed framework. A future relevant work would be to learn models with latent states and an unsupervised learning.	
\end{itemize}

\bibliography{hermes_paper}
\bibliographystyle{tmlr}

\end{document}